Giant Low Surface Brightness Galaxies As Merger Remnants

### 2016-08-01 glsb.tex
edit

XXX-in this paper we blah dee blah. In \S\ref{sec:obs} blah blah.

%\citet{Pickering97} has the \HI\ rotation curves of these guys.  For UGC 6614, has M \HI\ $= 2.5 \pm0.2\times 10^{10} \msun h_{75}^{-2}$.  And $3.6 \pm 0.4 \times 10^{10} \msun h_{75}^{-2}$for Malin 2.  Notes that these are galaxies that are massive and DM dominated.  In Malin 2, they find some gas moving relative to the regular rotating disk, and suspect that a dwarf fell though and stimulated star formation and this could be how stars managed to form in the low density environment.

@@

%From \citet{Matthews01b}, \HI\ measurements on the gals:
%name          rms   Fmax W20   W50   HRV  e_HRV S\HI\  eS\HI\  S/N
%F 568-6    *  2.29  17.5 420   393   13829  7  4.20  0.64  7.6 
%UGC 6614   *  2.86  81.2 290   255    6351  2 15.94  0.72 28.4 
%\citet{Matthews01b} also lists some incs values--calls Malin2 an Scd/Sd galaxy, has some good tables that are only available on-line.  I'll need to compare my stellar kinematics to the \HI\ kinematics.

@@

%\citet{Das10} find CO in Malin 2.  Make some points about it being stable

@@

\begin{deluxetable*}{l l c c c c c c}
%\tabletypesize{\small \footnotesize \scriptsize}
\tabletypesize{\small}
%\rotate
\tablewidth{0pt}
%\tablenum{num}
%\tablecolumns{num}
%\tableheadfrac{num}
\tablecaption{Basic Properties of GLSB Galaxies}
\tablehead{
\colhead{Name} & \colhead{SDSS ID}   &  \colhead{z}  &\colhead{$r$}  & \colhead{$R_e$} 
& \colhead{$\sigma$}  & \colhead{\HI\ Mass} & \colhead{Stellar Mass} \\
     &  \colhead{(DR 10)}   &     &  \colhead{(mag)}    & \colhead{arcsec} & \colhead{\kms} & \colhead{log$_{10}$M$_{\odot}$} & \colhead{log$_{10}$M$_{\odot}$}
\label{basic_prop}}    
\startdata
Malin 1 & 1237661068728139925 & 0.083 & 16.70 & 1.86 & 177 &  10.9 & 11.1 \\
Malin 2 & 1237667735572381852 &   0.046 & 14.34 & 7.17 & 202  &  10.6 & 11.3 \\
UGC 06614 & 1237668585435365393 &  0.021 & 13.52 & 7.3 & 150 &  10.5 & 10.9
\enddata
\tablecomments{\HI\ masses updated to our adopted cosmological parameters from \citet{Pickering97}.  The rest taken from SDSS. Stellar masses calculated from \citet{Conroy09}}
\end{deluxetable*}

% SDSS stellar masses.  using stellarMassFSPSGranEarlyDust:
% Malin 2:  logMass	11.3307373297902
% Malin 1:  logMass	11.1328074307929
% UGC 6614: logMass	10.8745260407613

% Updating the Pickering MI masses:
% M_\HI\ = 2.36e5 * D^2 * int(S)
% D = c(z+0.5(1.0-q_0)z^2)/H_0
% int(S) = 4.4 Jy km/s for Malin 2
%        = 15.0 for UGC 6614
% cosmolopy using h = 0.704.  
% cosmolopy says D_L = 202.9 and 90.9 Mpc
% so that gives me M_\HI\'s of 4.27e10 and 2.92e10

@@

\begin{deluxetable}{l c c }
\tabletypesize{\small}
\tablewidth{0pt}
\tablecaption{Fitted Structural Properties}
\tablehead{
\colhead{Name} & \colhead{$h_r$}   &  \colhead{inclination}    \\
     &  \colhead{(arcsec)}   & \colhead{(deg)}      
\label{fitted_params}}    
\startdata
Malin 2 & 6.1 & 35.5  \\
UGC 06614 & 5.7 & 33.4
\enddata
\end{deluxetable}

@@

%Observations were made using the Dual Imaging Spectrograph (DIS) on
%the ARC 3.5 meter telescope at Apache Point Observatory (APO).  Malin 2 was observed on 4/01/13, and UGC 06614 was observed 5/01/13.  Both galaxies were observed with the  gratings and multiple positions of the 1.5\arcsec\ slit.
%-insert table with galaxy name, position, total exposure time
%-insert figure showing slit positions overlayed on the SDSS thumbnails.

\section{Comparisons to Other Galaxies}

By eye, GLSB systems resemble large faint spiral disks at large radii, and regular ellipticals at their centers.  In the Sloan Digital Sky Survey Data Release 10 (SDSS DR10) \citep{Ahn14}, the Galaxy Zoo classifications \citep{Willett13} are spiral for Malin 1 and 2 and UGC 6614.  In this section, we compare the properties of the GLSB centers and disks with other systems.

@@

XXX--do find-replace check for Malin 2 and UGC 6614 to Malin~2 and UGC~6614

@@

The extracted rotation curves imply large dark matter halos.  For Malin 2, $V_{circ}\sim 280$ \kms at $R=42$ kpc implies a mass of $7.6\times 10^{11} \msun$ (compared to an SDSS stellar mass of $1.4\times 10^{11}$).  For UGC 6614, $V_{circ}\sim 200$ \kms at $R=17$ kpc gives a dynamical mass of $1.6\times 10^{11} \msun$ (compared to a stellar mass of $0.46\times 10^{11} \msun$ ).

Thus, using just the stellar kinematics, for Malin~2, we reach a radius of $\sim 6.2 R_e$, and find a dark matter fraction of 81\%.  For UGC 6614, we reach a radius of  $\sim 5.2 R_e$ and find a dark matter fraction of 71\%.

@@

%-crap, whatever happened to the gas kinematics?  There are emission lines out there--are they bright enough to get in single fibers of not?  Looks like for Malin 2 the emission lines are only coming from one side of the galaxy.  UGC 6614 is mostly emission from the center.

@@

\section{Star Formation Histories}

@@

Figures~\ref{malin2_binmap} and~\ref{ugc_binmap} show the fibers that were binned and the resulting spectra for Malin 2 and UGC 06614.  Both galaxies have elliptical-galaxy like spectra in their centers, with outskirts that have significant emission lines.  We find the best-fitting single model spectrum for each observed binned spectrum. When fitting, we include a low order multiplicative polynomial to match any dust extinction or flux calibration errors in the observations.  This means our fits are for the relative star formation histories, but not the absolute star formation rates (e.g., a model constant star formation history of 1 M$_\sun$ yr$^{-1}$ is an equally good fit as 10 M$_\sun$ yr$^{-1}$.)

@@

\section{Discussion}

@@

These are incredibly steep metallicity gradients.  The logarithmic gradients $\delta \rm{[Z/H]}\ \log{R/R_e}$ is $\sim -1.6$ for Malin 2 and $\sim -0.68$ for UGC 6614.  These are very steep compared to gradients typically observed in elliptical galaxies \citep{Greene12, Greene15}.  Other cites for elliptical gradients \citep{Kuntschner10}

@@

% malin 2: R_e = 7.17"
% UGC : 7.3"

@@

\section{Conclusions}

@@

\begin{figure}
\plottwo{Plots/sdssbinned_Malin2.eps}{Plots/Malin2mu.eps}
\epsscale{.5}
\plotone{Plots/Malin2profile_fit.eps}
\epsscale{1}
\caption{Upper Left:  SDSS image of Malin 2.  Upper Right:  Image made with the VIRUS-P datacube synthetic fiber magnitudes.  Lower:  An exponential fit to the fiber magnitudes.  Red points show fibers that were clipped while the solid line shows the best-fit.  \label{Malin2_fibers} }
\end{figure}

@@

\begin{figure}
\plottwo{Plots/UGC06614bin_map.eps}{Plots/UGC06614binned_spec.eps}
\caption{Left:  Map showing which fibers were binned for UGC 6614.  Right:  The binned spectra, color-coded to match the map on the left.  As with Malin~2, the inner spectra show strong metallicity absorption features while the outer spectra show emission lines. \label{ugc_binmap}}
\end{figure}

@@

Should probably figure out if these things are fast or slow rotators by the \citet{Emsell11} definition.  Also see \citet{Capp11} that says slow rotators are very rare in the field.  So it might be strange if Malin 2 and UGC 6614 are slow rotators.

>>>>

In \S\ref{sec:obs} we present new observations of Malin~2 and UGC~6614. In \S\ref{sec:compare_gals}, we compare the general properties of GLSB galaxies to sample of galaxies drawn from SDSS and the THINGS survey.  In \S\ref{sec:kinematics} and \S\ref{sec:sfh} we measure the stellar kinematics and star formation histories from our spectra. In \S\ref{sec:discussion}, we discuss what our observations imply for the possible formation mechanisms for GLSB systems.

%\citet{Pickering97} has the \HI\ rotation curves of these guys.  For UGC~6614, has M \HI\ $= 2.5 \pm0.2\times 10^{10} \msun h_{75}^{-2}$.  And $3.6 \pm 0.4 \times 10^{10} \msun h_{75}^{-2}$for Malin~2.  Notes that these are galaxies that are massive and DM dominated.  In Malin~2, they find some gas moving relative to the regular rotating disk, and suspect that a dwarf fell though and stimulated star formation and this could be how stars managed to form in the low density environment.

@@

%From \citet{Matthews01b}, \HI\ measurements on the gals:
%name          rms   Fmax W20   W50   HRV  e_HRV S\HI\  eS\HI\  S/N
%F 568-6    *  2.29  17.5 420   393   13829  7  4.20  0.64  7.6 
%UGC~6614   *  2.86  81.2 290   255    6351  2 15.94  0.72 28.4 
%\citet{Matthews01b} also lists some incs values--calls Malin2 an Scd/Sd galaxy, has some good tables that are only available on-line.  I'll need to compare my stellar kinematics to the \HI\ kinematics.

@@

%\citet{Das10} find CO in Malin~2.  Make some points about it being stable

@@

\begin{deluxetable*}{l l c c c c c c}
%\tabletypesize{\small \footnotesize \scriptsize}
\tabletypesize{\small}
%\rotate
\tablewidth{0pt}
%\tablenum{num}
%\tablecolumns{num}
%\tableheadfrac{num}
\tablecaption{Basic Properties of GLSB Galaxies}
\tablehead{
\colhead{Name} & \colhead{SDSS ID}   &  \colhead{z}  &\colhead{$r$}  & \colhead{$R_e$} 
& \colhead{$\sigma$}  & \colhead{\HI\ Mass} & \colhead{Stellar Mass} \\
     &  \colhead{(DR 10)}   &     &  \colhead{(mag)}    & \colhead{arcsec} & \colhead{\kms} & \colhead{log$_{10}$M$_{\odot}$} & \colhead{log$_{10}$M$_{\odot}$}
\label{basic_prop}}    
\startdata
Malin 1 & 1237661068728139925 & 0.083 & 16.70 & 1.86 & 177 &  10.9 & 11.1 \\
Malin~2 & 1237667735572381852 &   0.046 & 14.34 & 7.17 & 202  &  10.6 & 11.3 \\
UGC 06614 & 1237668585435365393 &  0.021 & 13.52 & 7.3 & 150 &  10.5 & 10.9
\enddata
\tablecomments{\HI\ masses updated to our adopted cosmological parameters from \citet{Pickering97}.  The rest taken from SDSS. Stellar masses calculated from \citet{Conroy09}}
\end{deluxetable*}

% SDSS stellar masses.  using stellarMassFSPSGranEarlyDust:
% Malin~2:  logMass	11.3307373297902
% Malin 1:  logMass	11.1328074307929
% UGC~6614: logMass	10.8745260407613

% Updating the Pickering MI masses:
% M_\HI\ = 2.36e5 * D^2 * int(S)
% D = c(z+0.5(1.0-q_0)z^2)/H_0
% int(S) = 4.4 Jy km/s for Malin~2
%        = 15.0 for UGC~6614
% cosmolopy using h = 0.704.  
% cosmolopy says D_L = 202.9 and 90.9 Mpc
% so that gives me M_\HI\'s of 4.27e10 and 2.92e10

@@

\begin{deluxetable}{l c c }
\tabletypesize{\small}
\tablewidth{0pt}
\tablecaption{Fitted Structural Properties}
\tablehead{
\colhead{Name} & \colhead{$h_r$}   &  \colhead{inclination}    \\
     &  \colhead{(arcsec)}   & \colhead{(deg)}      
\label{fitted_params}}    
\startdata
Malin~2 & 6.1 & 35.5  \\
UGC 06614 & 5.7 & 33.4
\enddata
\end{deluxetable}

@@

%Observations were made using the Dual Imaging Spectrograph (DIS) on
%the ARC 3.5 meter telescope at Apache Point Observatory (APO).  Malin~2 was observed on 4/01/13, and UGC 06614 was observed 5/01/13.  Both galaxies were observed with the  gratings and multiple positions of the 1.5\arcsec\ slit.
%-insert table with galaxy name, position, total exposure time
%-insert figure showing slit positions overlayed on the SDSS thumbnails.

\section{Comparisons to Other Galaxies}\label{sec:compare_gals}

By eye, GLSB systems resemble large faint spiral disks at large radii, and regular ellipticals at their centers.  In the Sloan Digital Sky Survey Data Release 10 (SDSS DR10) \citep{Ahn14}, the Galaxy Zoo classifications \citep{Willett13} are spiral for Malin 1 and 2 and UGC~6614.  In this section, we compare the properties of the GLSB centers and disks with other systems.

@@

@@

The extracted rotation curves imply large dark matter halos.  For Malin~2, $V_{circ}\sim 280$ \kms at $R=42$ kpc implies a mass of $7.6\times 10^{11} \msun$ (compared to an SDSS stellar mass of $1.4\times 10^{11}$).  For UGC~6614, $V_{circ}\sim 200$ \kms at $R=17$ kpc gives a dynamical mass of $1.6\times 10^{11} \msun$ (compared to a stellar mass of $0.46\times 10^{11} \msun$ ).

Thus, using just the stellar kinematics, for Malin~2, we reach a radius of $\sim 6.2 R_e$, and find a dark matter fraction of 81\%.  For UGC~6614, we reach a radius of  $\sim 5.2 R_e$ and find a dark matter fraction of 71\%.

@@

%-crap, whatever happened to the gas kinematics?  There are emission lines out there--are they bright enough to get in single fibers of not?  Looks like for Malin~2 the emission lines are only coming from one side of the galaxy.  UGC~6614 is mostly emission from the center.

@@

\section{Star Formation Histories}\label{sec:sfh}

@@

Figures~\ref{malin2_binmap} and~\ref{ugc_binmap} show the fibers that were binned and the resulting spectra for Malin~2 and UGC 06614.  Both galaxies have elliptical-galaxy like spectra in their centers, with outskirts that have significant emission lines.  We find the best-fitting single model spectrum for each observed binned spectrum. When fitting, we include a low order multiplicative polynomial to match any dust extinction or flux calibration errors in the observations.  This means our fits are for the relative star formation histories, but not the absolute star formation rates (e.g., a model constant star formation history of 1 M$_\sun$ yr$^{-1}$ is an equally good fit as 10 M$_\sun$ yr$^{-1}$.)

@@

\section{Discussion}\label{sec:discussion}

@@

These are incredibly steep metallicity gradients.  The logarithmic gradients $\delta \rm{[Z/H]}\ \log{R/R_e}$ is $\sim -1.6$ for Malin~2 and $\sim -0.68$ for UGC~6614.  These are very steep compared to gradients typically observed in elliptical galaxies \citep{Greene12, Greene15}.  Other cites for elliptical gradients \citep{Kuntschner10}

@@

% Malin~2: R_e = 7.17"
% UGC : 7.3"

@@

\section{Conclusions}\label{sec:conclusions}

@@

\begin{figure}
\plottwo{Plots/sdssbinned_Malin2.eps}{Plots/Malin2mu.eps}
\epsscale{.5}
\plotone{Plots/Malin2profile_fit.eps}
\epsscale{1}
\caption{Upper Left:  SDSS image of Malin~2.  Upper Right:  Image made with the VIRUS-P datacube synthetic fiber magnitudes.  Lower:  An exponential fit to the fiber magnitudes.  Red points show fibers that were clipped while the solid line shows the best-fit.  \label{Malin2_fibers} }
\end{figure}

@@

\begin{figure}
\plottwo{Plots/UGC06614bin_map.eps}{Plots/UGC06614binned_spec.eps}
\caption{Left:  Map showing which fibers were binned for UGC~6614.  Right:  The binned spectra, color-coded to match the map on the left.  As with Malin~2, the inner spectra show strong metallicity absorption features while the outer spectra show emission lines. \label{ugc_binmap}}
\end{figure}

@@

Should probably figure out if these things are fast or slow rotators by the \citet{Emsell11} definition.  Also see \citet{Capp11} that says slow rotators are very rare in the field.  So it might be strange if Malin~2 and UGC~6614 are slow rotators.
